### 2010-11-15 paper/tables/sha3_flags.tex
Expanded chosen candidates section

\begin{table}
  \centering
  \begin{tabular}{ | l | l | }
    \hline
    \textbf{candidate} & \textbf{from}  \\ \hline
     blake       & mssse3 -m32 -march=native -mtune=native -O -fomit-frame-pointer  \\ \hline
     bmw         & -funroll-loops -m32 -march=pentium4 -O3 -fomit-frame-pointer     \\ \hline
     cubehash    & none                                                             \\ \hline
     echo        & -funroll-loops -m32 -march?prescott -O2 -fomit-frame-pointer     \\ \hline
     fugue       & -funroll-loops -O -fomit-frame-pointer                           \\ \hline
     groestl     & -m32 -march=pentium3 -O -fomit-frame-pointer                     \\ \hline
     hamsi       & -msse2 -funroll-loops -m32 -march=pentium-m -O3 -fomit-frame-pointer\\ \hline
     jh          & none                                                             \\ \hline
     keccak      & -msse2 -march=barcelona -O3 -fomit-frame-pointer                 \\ \hline
     luffa       & -msse2 -funroll-loops -march=k8 -O3 -fomit-frame-pointer         \\ \hline
     shabal      & -funroll-loops -m32 -march=pentium2 -O -fomit-frame-pointer      \\ \hline
     shavite3    & -funroll-loops -march=i486 -O -fomit-frame-pointer               \\ \hline
     simd        & -m32 -march=barcelona -O2 -fomit-frame-pointer                   \\ \hline
     skein       & -march=pentium2 -O2 -fomit-frame-pointer                         \\ \hline
  \end{tabular}
  \caption{\ac{GCC} compiler flags used for the implementations}
  \label{tbl:sha:compilerflags}
\end{table}