### 2010-12-11 paper/tables/supercop_sha256d.tex
Changed supercop double sha256 to worst-case scenario

\begin{table}
  \centering
  \begin{tabular}{ | c | c | }
    \hline
    \textbf{Size} & \textbf{Median} \\ \hline
    8    & 898.1 \\ \hline
    64   & 127.93 \\ \hline
    576  & 53.85 \\ \hline
    1536 & 20.19 \\ \hline
    4096 & 17.23 \\ \hline
  \end{tabular}
  \caption{Estimates for double SHA-256}
  \label{tbl:supercop:sha256d}
\end{table}