### 2010-12-11 paper/tables/sha3_flags.tex
Fixed compiler flags

\begin{table}
  \centering
  \begin{tabular}{ | l | l | }
    \hline
    \textbf{Candidate} & \textbf{Compiler flags}            \\ \hline
     BLAKE       & -m32 -march=native -mtune=native -O      \\ \hline
     \ac{BMW}    & -funroll-loops -m32 -march=pentium4 -O3  \\ \hline
     Cubehash    & -O2                                      \\ \hline
     Echo        & -funroll-loops -m32 -march=prescott -O2  \\ \hline
     Fugue       & -funroll-loops -O                        \\ \hline
     Grøstl      & -m32 -march=pentium3 -O                  \\ \hline
     Hamsi       & -funroll-loops -m32 -march=pentium-m -O3 \\ \hline
     JH          & -m32 -march=core2 -O                     \\ \hline
     Keccak      & -march=barcelona -O3                     \\ \hline
     Luffa       & -funroll-loops -march=k8 -O3             \\ \hline
     Shabal      & -funroll-loops -m32 -march=pentium2      \\ \hline
     Shavite3    & -funroll-loops -march=i486 -O            \\ \hline
     Simd        & -m32 -march=barcelona -O2                \\ \hline
     Skein       & -march=pentium2 -O2                      \\ \hline
  \end{tabular}
  \caption{GCC compiler flags used on the implementations}
  \label{tbl:sha3:compilerflags}
\end{table}