### 2010-12-14 paper/tables/supercop_sha256d.tex
Added more calculations to sha256d estimate

\begin{table}
  \centering
  \begin{tabular}{ | c | c | c | c |}
    \hline
    \textbf{Size} & \textbf{Worst case} & \textbf{Interpolation} & \textbf{Best case} \\ \hline
    8    & 898.1  & 655.53  & 332.10 \\ \hline
    64   & 127.93 & 97.61  & 57.18 \\ \hline
    576  & 53.85  & 24.66  & 20.17 \\ \hline
    1536 & 20.19  & 18.89  & 17.23 \\ \hline
    4096 & 17.23  & 16.76  & 16.13 \\ \hline
  \end{tabular}
  \caption{Estimates for medians using double \ac{SHA}-256}
  \label{tbl:supercop:sha256d}
\end{table}